### 2019-03-23 R/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
Upload and cleanup 23/3/2019

# -----------------------------------------------------------------------------------------------
# Analyse changes in pelagic stocks.Rmd
#
# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# 13/01/2019 now only figures and tables
# 15/01/2019 added scale jump figure with explanatory variables
# -----------------------------------------------------------------------------------------------

@@

# Libraries
library(tidyverse) # for piping and easy coding
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(RColorBrewer) # colours
library(lubridate)
library(scales)
library(directlabels)  # for printing labels at end of geom lines

@@

  # benchmarks
  left_join(d.benchmarks, by=c("stock", "stockkeylabelold", "assessmentyear")) %>% 
  mutate(benchmark = ifelse(is.na(benchmark), 0, benchmark))

@@

# Calculations of scale-differences and scale jumps ==========================================

@@

# data series with most recent (last) estimation
d.last <-
  d.pairs %>% 
  select(-assessmentyearprior) %>% 
  left_join(d.selection, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
  mutate(label="last") %>% 
  rename(assessmentyear = assessmentyearlast)

# data series with previous (prior) estimation
d.prior <-
  d.pairs %>% 
  select(-assessmentyearlast) %>% 
  left_join(d.selection, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
  mutate(label="prior") %>% 
  rename(assessmentyear = assessmentyearprior)

# pairwise data roughly
d.pairwise.roughly <-
  bind_rows(d.last, d.prior) %>% 
  group_by(stock, assessmentpair, variable) %>%

@@

@@

@@

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

myvar     <- "stocksize"
myspecies <- "mac"
firstyear <- 1991

@@

  d.pairwise %>%

@@

  geom_line(data=filter(d.pairwise.roughly, 
                        variable == myvar, 
                        grepl(myspecies, stockkeylabelold),
                        as.numeric(substr(assessmentpair,1,4)) >= firstyear), size=0.5) +

@@

<!-- ################################################################################## -->
<!-- Plot scale jumps                                                                   -->
<!-- ################################################################################## -->

# plot scale jumps between assessments
d.jumps %>% 
  distinct(stock, variable, assessmentyear, phi) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

@@

  geom_point(data=filter(d.assessments, benchmark == 1),
            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +

@@

@@

d.jumps %>% 
  distinct(stock, variable, assessmentyear, phi) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

@@

  geom_point(data=filter(d.assessments, benchmark == 1),
            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +

@@

@@

  geom_point(data=filter(d.assessments, benchmark == 1),
            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +

@@

>>>>

# -----------------------------------------------------------------------------------------------
# Analyse changes in pelagic stocks.Rmd
#
# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# 13/01/2019 now only figures and tables
# 15/01/2019 added scale jump figure with explanatory variables
# 22/02/2019 focussed on Pairwise Comparison of Means = scale jumps under different conditions
# -----------------------------------------------------------------------------------------------

@@

# Libraries
library(tidyverse) # for piping and easy coding
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(ggExtra)   # for specific type of plots e.g. ggMarginal
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(RColorBrewer) # colours
library(lubridate)
library(scales)
library(directlabels)  # for printing labels at end of geom lines

@@

  # benchmarks
  full_join(d.benchmarks, by=c("stock", "stockkeylabelold", "assessmentyear")) %>% 
  # left_join(d.benchmarks, by=c("stock", "stockkeylabelold", "assessmentyear")) %>% 
  mutate(benchmark = ifelse(is.na(benchmark), 0, benchmark))

@@

# Calculations of Pairwise Comparison of Mean and scale jumps ==========================================

pairwise_comparisons <- function(removelastyears=0, numberofyears=10) {

  # data series with most recent (last) estimation of the pair
  d.pairs.last <-
    d.pairs %>% 
    select(-assessmentyearprior) %>% 
    left_join(d.selection, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
    mutate(label="last") %>% 
    rename(assessmentyear = assessmentyearlast)

  # data series with previous (prior) estimation of the pair
  d.pairs.prior <-
    d.pairs %>% 
    select(-assessmentyearlast) %>% 
    left_join(d.selection, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
    mutate(label="prior") %>% 
    rename(assessmentyear = assessmentyearprior)

  # pairwise data roughly
  d.pairwise.roughly <-
    bind_rows(d.pairs.last, d.pairs.prior) %>% 
    group_by(stock, assessmentpair, variable) %>%

  d.pcm <-
    d.pairwise %>%

  res <- list("d.pcm"= d.pcm, 
              "d.pairwise.roughly"=d.pairwise.roughly,
              "d.pairwise" =d.pairwise)

  return(res)

} # End of pairwise_comparisons function

tmp <- pairwise_comparisons(removelastyears = 0, numberofyears=5)

# REMOVE FROM HERE !!

@@

# data series with most recent (last) estimation of the pair
d.pairs.last <-
  d.pairs %>% 
  select(-assessmentyearprior) %>% 
  left_join(d.selection, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
  mutate(label="last") %>% 
  rename(assessmentyear = assessmentyearlast)

# data series with previous (prior) estimation of the pair
d.pairs.prior <-
  d.pairs %>% 
  select(-assessmentyearlast) %>% 
  left_join(d.selection, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
  mutate(label="prior") %>% 
  rename(assessmentyear = assessmentyearprior)

# pairwise data roughly
d.pairwise.roughly <-
  bind_rows(d.pairs.last, d.pairs.prior) %>% 
  group_by(stock, assessmentpair, variable) %>%

@@

# CONTINUE FROM HERE

@@

```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}

## NOT SHOWN

@@

res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)

myvar     <- "stocksize"
myspecies <- "mac"
firstyear <- 2008

@@

  res$d.pairwise %>%

@@

  geom_line(data=filter(res$d.pairwise.roughly, 
                        variable == myvar, 
                        grepl(myspecies, stockkeylabelold),
                        as.numeric(substr(assessmentpair,1,4)) >= firstyear), size=0.5) +

@@

<!-- ################################################################################## -->
<!-- Plot pairwise comparison means                                                     -->
<!-- ################################################################################## -->

```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}

res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)

threshold <- 0.10

# plot scale jumps between assessments
res$d.pcm %>% 
  distinct(stock, variable, assessmentyear, phi) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

@@

  geom_line (aes(colour=factor(decade))) +
  geom_point(aes(colour=factor(decade), size = ifelse( abs(value) > threshold, 0.3, 0.0001))) +
  # geom_point(aes(colour=factor(decade), size = abs(value) )) +

  geom_point(data=filter(d.benchmarks, benchmark == 1),
            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +

@@

```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}

## NOT SHOWN

res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)

threshold <- 0.10

@@

res$d.pcm %>% 
  distinct(stock, variable, assessmentyear, phi) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

@@

  geom_point(data=filter(d.benchmarks, benchmark == 1),
            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +

@@

res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)

threshold <- 0.10

@@

  geom_point(data=filter(d.benchmarks, benchmark == 1),
            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +

  # coord_cartesian(xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
  coord_cartesian(ylim=c(-1, 1),
                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded

  # ggMarginal(data=d.assessments, type="histogram")
  scale_fill_manual(values = myBlues) +

@@

# filter(d.assessments, benchmark == 1) %>% View()
